---
title: "esm 206 lab 5"
author: "Haley Grant"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Objectives:

- Getting counts for different groups
- Use {lubridate} to parse dates
- Find confidence intervals & do t-tests with t.test()
- Heatmap with geom_tile()

```{r, include = FALSE}

# Attach packages

library(tidyverse)
library(here)
library(janitor)

```

```{r}

# Import lobster data

lobster_abundance <- read_csv(here::here("data","lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()

```

Use the tidyr::uncount() function to expand a frequency variable

```{r}

lobster_tidy <- lobster_abundance %>%
  tidyr::uncount(lobster_count)
  
```

Yaya! Now each lobster has its own row. This is tidy format. 

Let's look at this a bunch of ways:

```{r}

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site),
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
  facet_wrap(~site, scales = "free")

```

#### 5. Parse dates for lobster observations

Notice that the existing class of the `date` variable is a character. Boo.
```{r}
class(lobster_abundance$date)
```

The `lubridate` package (part of the `tidyverse`), and is built to make it easier to deal with date-time data. Here, we'll use `lubridate::mdy()` to help R understand it's a date, and work with the pieces (month, day and year) more easily. 

**Note**: we're using `mdy()` because that's the existing format of the date column (mm/dd/yy). Check `?ymd` to see all the different options based on the format of your date column. Cool. 

Add a new column with `mutate()` that is a *date* using `mdy()`: 
```{r}
lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
```

Check it out! The *date_new* column is in nice ISO date format. Let's check the class:
```{r}
class(lobster_date$date_new)
```

Wooooo.

Now that it's in date format, we can parse it. See `?month` and `?year` - to get or set components of a date or date-time. So here, we'll use `lubridate::month()` and `lubridate::year()` to create separate columns for the month and year. 

For `month()`, we can even automatically convert to month abbreviation with argument `label = TRUE`. 

```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(obs_month = lubridate::month(date_new, label = TRUE),
         obs_year = lubridate::year(date_new))
```

Count lobsters by different groupings...

Let's say I want to count the number of lobsters by year and month

```{r}
lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

lobster_ym

lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)

lobster_ysite

lobster_site <- lobster_parse_date %>% 
  count(site)

lobster_site

# group_by + summarize + n()
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n()
  )

```

Confidence intervals and t-tests

Use the 't-test()' function to find confidence intervals and perform t-tests

```{r}

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

my_vector <- c(3, 5, 9, 15, 24)

t.test(ivee_lobsters)

```

Two-sample t-test:
Is there a significant difference in mean lobster lengths between Naples Reef and Mohawk Reef?

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample, mohk_sample)

mohk_napl_ttest


```

Mean lobster size differed significantly between Mohawk and Naples reefs

Maybe an easier way...
```{r}
lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL","MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)

ttest_2
```

Make a geom_tile heatmap

```{r}
ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

